Using temperature distributions of active regions to investigate flare activity

### 2015-03-17 paper.tex
Minor edits based on Huw's suggestions

Studies of coronal temperatures date back several decades. % Stick some references here to early DEM studies
However, the Atmospheric Imaging Assembly (AIA; \cite{Lemen2011}) which began observing the Sun in 2010 allows unprecedented access to observations with very high spatial and temporal resolution.
This allos us to investigate temperature distributions on small spatial scales with high cadence. % Don't much like this sentence, but oh well
The six Fe-based EUV channels of AIA also provide greater constraints on temperature solutions than previous coronal imagers \citep{Guennou2012, Guennou2012a}.

This work makes use of the capabilities of AIA in order to investigate the temperature distributions of several flaring active regions.
The aim is to search for a common signature of flare activity in these temperature distributions before flares occur, which, if found, could form the basis of a flare prediction algorithm.
To achieve this, we look at temperature distributions of the active regions over the 30 minutes before a flare occurs, and we also directly compare the temperatures of the corresponding active regions at several different times before the flares.

@@

%===========================================================================
\section{Method}
\subsection{Temperature analysis}
The core of this analysis is the temperature map method described by \cite{Leonard} (henceforth Paper I). \defcitealias{Leonard}{Paper I}
Briefly, this method compares the relative brightnesses in each AIA wavelength channel at a given time and infers the best-fitting temperature from that comparison. % This can be phrased more descriptively
This allows the temperature of the corona to be estimated for every pixel of an AIA image, resulting in very high resolution temperature maps.
Crucially, this method produces temperature maps extremely quickly, which allows us to investigate coronal temperatures during localised dynamic events such as flares.
With other, slower methods, such studies are infeasible due to the computation time required.
An example temperature map is shown in Figure \ref{fig:example_tmap}.

The \citetalias{Leonard} method also allows easy `tracking' of solar regions by cropping the temperature map around a given pair of Carrington heliographic coordinates, which rotate at approximately the rate of solar rotation.
This tracking is demonstrated in Figures \ref{fig:trackdemo1} and \ref{fig:trackdemo2}, each of which shows a cropped region of an AIA 17.1nm image containing active region AR11158, the corresponding temperature map and a larger 17.1nm image showing the location of the region in the corona.

@@

% Put this table in an appendix if possible
% Also try and put two sets of columns on a page so it takes up less space
\begin{longtable}{c|c|c}
	\caption{Start times and associated active regions of solar flares studied during 2011-02 and 2011-03.}\\
		Date and time & GOES class & NOAA Active region \\
		\hline
		\input{tempplotsmean/flarelist.txt}
	\label{tab:flares}
\end{longtable}

@@

%===========================================================================
\section{Discussion and conclusions}
From this study, it appears that there is no clear link between solar flares and coronal temperatures.
This is probably mostly due to the relatively small number of flares studied - a much larger sample size would have to be used to properly determine any link or lack thereof.
In particular, this study only included a single X class flare, which as the most energetic are the kind we would most wish to predict.
Future works should therefore aim to consider more large flares.
Non-flaring active regions should also be taken into account and compared against the results presented here.
However, this work does demonstrate that it is now possible to study temperature distributions of the corona in high resolution and on short timescales using fast temperature analysis tools such as the one described in \citetalias{Leonard}.

@@

%===========================================================================
\begin{acknowledgements}
	This research has made use of SunPy, an open-source and free community-developed solar data analysis package written in Python \citep{Mumford2013}.
\end{acknowledgements}

>>>>

Studies of coronal temperatures date back several decades. % Stick some references here to early DEM studies
However, the Atmospheric Imaging Assembly (AIA; \cite{Lemen2011}) which began observing the Sun in 2010 allows unprecedented access to observations with very high spatial and temporal resolution.
This allows investigation of temperature distributions on small spatial scales with high cadence. % Don't much like this sentence, but oh well
The six Fe-based EUV channels of AIA also provide greater constraints on temperature than previous coronal imagers \citep{Guennou2012, Guennou2012a}.

This work makes use of the capabilities of AIA in order to investigate the temperature distributions of several flaring active regions.
The aim is to search for a common signature of flare activity in these temperature distributions before flares occur, which, if found, could form the basis of a flare prediction algorithm.

%===========================================================================
\section{Method}
This investigation looks at temperature distributions of active regions over the 30 minutes before a flare occurs.
We also directly compare the temperatures of the corresponding active regions at several different times before the flares.

@@

\subsection{Temperature analysis}
The core of this analysis is the temperature map method described by \cite{Leonard} (henceforth Paper I). \defcitealias{Leonard}{Paper I}
Briefly, this method compares the relative brightnesses in each AIA wavelength channel at a given time and infers the best-fitting temperature from that comparison. % This can be phrased more descriptively
This allows the temperature of the corona to be estimated for every pixel of an AIA image, resulting in very high resolution temperature maps.
Crucially, this method produces temperature maps extremely quickly, which allows us to investigate coronal temperatures during localised dynamic events such as flares.
With other, slower methods, such studies are infeasible due to the computation time required.
An example temperature map is shown in Figure \ref{fig:example_tmap}.

The \citetalias{Leonard} method also allows easy `tracking' of solar regions by cropping the temperature map around a given pair of Carrington heliographic coordinates, which rotate at approximately the rate of solar rotation.
This tracking is demonstrated in Figures \ref{fig:trackdemo1} and \ref{fig:trackdemo2}, each of which shows a cropped region of an AIA 17.1nm image containing active region AR11158, the corresponding temperature map and a larger 17.1nm image showing the location of the region in the corona.
Note that the small, bright loops in the centre of the active region are highlighted in red, indicating temperatures of up to $log(T) \sim 6.3$, while some larger, fainter loops around the edges are at temperatures of around $log(T) \sim 6.15$.
These temperatures are consistant with values found by other authors for active region loops. % Reference the shit out of that comment.

@@

% Put this table in an appendix if possible
% Also try and put two sets of columns on a page so it takes up less space
\begin{longtable}{c|c|c}
	\caption{Start times and associated active regions of solar flares studied between 2011/02/01 and 2011/03/31.}\\
		Date and time & GOES class & NOAA Active region \\
		\hline
		\input{tempplotsmean/flarelist.txt}
	\label{tab:flares}
\end{longtable}

@@

%===========================================================================
\section{Discussion and conclusions}
From this study, it appears that there is no clear link between solar flares and the evolution of active region temperatures prior to flaring.
This is probably mostly due to the relatively small number of flares studied - a much larger sample size would have to be used to properly determine any link or lack thereof.
In particular, this study only included a single X class flare, which are the most energetic and the kind we would most wish to predict.
Future works should therefore aim to consider larger flares.
Non-flaring active regions should also be taken into account and compared against the results presented here.
However, this work does demonstrate that it is now possible to study temperature distributions of the corona in high resolution and on short timescales using fast temperature analysis tools such as the one described in \citetalias{Leonard}.

@@

%===========================================================================
\begin{acknowledgements}
	This research has made use of SunPy, an open-source and free community-developed solar data analysis package written in Python \citep{Mumford2013}.
	AIA images courtesy of NASA/SDO and the AIA science team.
\end{acknowledgements}
